---
title: 'Bellabeat Case Study: Smart Device Usage Analysis'
author: "suhail"
date: "2026-01-21"
output: html_document
---


will produce a repo with the following deliverables:                            [DATE :- 18-01-2026]

1.	A clear summary of the business task 
2.	 A description of all data sources used
3.	 Documentation of any cleaning or manipulation of data 
4.	 A summary of your analysis 
5.	 Supporting visualizations and key findings
6.	 Your top high-level content recommendations based on your analysis

### ASK PHASE 
#### Business Task :- 
“ Analyze smart-device usage patterns from available fitness tracker data to identify trends and insights that can guide Bellabeat’s product development and marketing strategy for non-smart wellness products. “
                                              
#### Stake holders :- 
  1. UrškaSršen : Bellabeat’s cofounder and Chief Creative Office                  2. SandoMur:  Mathematician and Bellabeat’s cofounder
  
### PREPARE
#### DATA IMPORTED FROM :-
1.	Public data_set :-  fitbit fitness tracker data                     🔗(https://www.kaggle.com/datasets/arashnic/fitbit)
2. Combined daily_activity and sleep-day data for good understanding 

#### importing data 
```{r}
fitbase_1 <- read.csv("~/coursera case study project/case stdy 2 _f/mturkfitbit_export_3.12.16-4.11.16/Fitabase Data 3.12.16-4.11.16/dailyActivity_merged.csv")

fitbase_2 <- read.csv("~/coursera case study project/case stdy 2 _f/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")

colnames(fitbase_1) 

```
####  arranging data 

```{r}
library(dplyr)
library(lubridate)

fitbase_1 <- fitbase_1 %>%
  mutate(ActivityDate = as.Date(ActivityDate, format = "%m/%d/%Y"))

fitbase_2 <- fitbase_2 %>%
  mutate(SleepDay = as.Date(SleepDay, format = "%m/%d/%Y %I:%M:%S %p")) %>%
  rename(ActivityDate = SleepDay)

fitbase <- left_join(fitbase_1, fitbase_2, by = c("Id", "ActivityDate"))

glimpse(fitbase)
```
	
	Glimps in r programming  👇
```{r}
glimpse(fitbase)
```

   

### PREPARE
Date : 19-1-26`
▼	 Tools Used `
1.	My SQL
2.	R (Programming)
3.	Tableau
4.	Excel 

      ▼Dividing process into two steps  :- 
      
                                                                🔹 Part 1: Activity-based behavior (ALL USERS)
                                                                •	Steps
                                                                •	Calories
                                                                •	Active minutes
                                                                •	Sedentary time                                                      🔹  Part 2: Sleep-based behavior (ONLY USERS WITH SLEEP DATA)
                                                                •	Filter users/days where sleep is available
                                                                •	Explicitly mention this limitation
                                                                 19-01-2026 
#### sorting fitbase using arrange()function

sorted the tables
                                                                
#### duplicate checks      
```{r}
library(dplyr)
dup_check <- fitbase %>%  count(Id,ActivityDate) %>% filter(n >1)
dup_check

```
####  Data Validation → checking data types 
```{r}
str(fitbase)

```
 21-01-2026 [ Analyzis using R ]
```{r}
library(dplyr)
user_summary <- fitbase %>% group_by(Id) %>% 
  summarise(avg_steps = mean(TotalSteps ,na.rm = TRUE),
            avg_calories = mean(Calories , na.rm = TRUE),
            avg_sedimentary_minutes = mean(SedentaryMinutes , na.rm = TRUE),
            avg_active_minutes = mean(VeryActiveMinutes +FairlyActiveMinutes +LightlyActiveMinutes))

str(user_summary)
View(user_summary)
```
###  Visualization
```{r}
library(ggplot2)

```
A. Average steps per user
```{r}

ggplot(user_summary,
       aes(x = factor(Id), y = avg_steps )) +
  geom_col() +
  labs(
    title = "Average Daily Steps per User",
    x = "User ID",
    y = "Average Steps"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```
  B. Sedimentary time vs steps 
```{r}
ggplot(user_summary, aes(x = avg_steps, y = avg_sedimentary_minutes)) +
  geom_point() +
  labs(
    title = "Sedimentary Time vs Average Steps",
    x = "Average Steps",
    y = "Average Sedentary Minutes"
  )
```
C. Calories Vs Activity
```{r}
ggplot(user_summary, aes(x = avg_active_minutes, y = avg_calories)) +
  geom_point() +
  labs(
    title = "Active Minutes vs Calories Burned",
    x = "Average Active Minutes",
    y = "Average Calories"
  )
```

### 🚀 BUSINESS RECOMMENDATIONS

✅ Recommendation 1: Habit-building over performance
Since most users show moderate activity with high sedentary time, Bellabeat should emphasize daily habit formation rather than intense fitness goals.

✅ Recommendation 2: Non-smart product positioning
  Non-smart wellness products (journals, hydration reminders, mindfulness tools) should target sedentary users, positioning wellness as achievable without technology overload.
  
✅ Recommendation 3: Marketing messaging
Campaigns should focus on “small daily progress”, aligning with users who are not highly active but consistently engaged.

✅ Recommendation 4: Sleep features (future opportunity)
Given the limited sleep data, Bellabeat can differentiate by encouraging manual sleep reflection in non-smart products to complement wearable insights.                                         